Make any stupid plot you want

### 2021-02-14 draft/sections/next.tex
0.0.1 of what have we done done

%"This chapter steps back from the details and considers the larger issues that have been raised. In particular, it considers how this research helps to accomplish the motivating goal of this dissertation, which is the development of computer-based graphical communication." - Mackinlay
"We have a unified mathematical description of understanding visualization.  Lots of people have worked on this and each has identified and described important aspects, but each has specialized in some way and it is not clear how to align them.  We can understand all of the prior work via our framework that pulls from {established math}"
that way you make clear what your contribution is, gives you a context to bring up all of the citations, a frame to contrast them, and then a nice lead into the math
 "we validate the framework by using it to re-design artist and data access layer of Maptlotlib, a general purpose visualization tool"

@@

This work generalizes previous research constraining visual encodings from data components to graphic components as equivariant maps to components that are N-dimensional. Furthermore, it precisely defines the glyph as the visual element constructed from data on a simplex in a simplacial complex where the simplex is discrete or continuous. This is a restatement of for example Bertin's definition of a line that encapsulates all points on the continuous line.  By modeling the data topology along with its variablity, this work also provides  a generalization of topology preservation as a deformation retraction from graphic space to data space. By using a functional paradigm, we can deconstruct the graphic to the glyph associated with each data point or even to pieces of a glyph; therefore the renderer has full  flexibility in how to to generate the image, while the graphic to data topology maps provide a book keeping mechanism to keep track of which part of the image belongs to which data point.

 \begin{itemize}
    \item prototype for scatter, line, bar
    \item simple composition into multibar
    \item commonly used container - dataframe
 \end{itemize}

\subsection{What are the limitations?}
%%%An interesting control problem will be how to keep the text and graphics consistent. If the rendering exchanges the horizontal and vertical axes of a bar chart, the text must also exchange its references to the axes.
\begin{itemize}
\item is at arc level
\item is at one dataset level - needs invariants/composites
\item static (what happens with interactive - but this is just update )
\item is compose(encodes) and ggplot is plot(encodes, glyph)
\item can't handle figerative/semantic equivariance (this actually could maybe be done as extension of fber)
\item explicitely does not care about effectiveness - is domain specific 
\item is only developed against Matplotlib
\item is written against AGG/SVG
\item layout
\item text
\item same as APT, is at gd level so cannot control render 
\item keeping all data in sync if transform is changed in one place (so on multi composed graphs, changing x on does it propopgate - part of Bertin invariance)
\end{itemize}
\subsection{What else needs to be implemented?}
\begin{itemize}
    \item heatmaps/ proof of concept of face simplex
    \item composition algebra
    \item boxplot - multiglyph composition
    \item debatable: non-trivial bundle like data on mobius backend
    \item interactivity 
    \item concurrent artist
    \item numpy wrapper
\end{itemize}
\begin{itemize}
    \item provisional implementation using new architecture in Matplotlib
    \item developer-aimed documentation (API + narrative)
    \item proof of concept 3rd party user-facing library with a biology partner
    \item academic paper about building visualization tools on top of the core architecture
\end{itemize}
Other things that may or may not play a big role would be working out the details of the S->K interaction for dashboards. By this I particularly mean when S is a composite object spanning all the figures shown in the dashboard and a component of K may have different compoents of S mapping into it slicing it in different ways.
\subsection{What are the lessons learned?}

>>>>

%"This chapter steps back from the details and considers the larger issues that have been raised. In particular, it considers how this research helps to accomplish the motivating goal of this dissertation, which is the development of computer-based graphical communication." - Mackinlay

@@

This work generalizes previous research constraining visual encodings from data components to graphic components as equivariant maps to components that are N-dimensional. Furthermore, it precisely defines the glyph as the visual element constructed from data on a simplex in a simplacial complex where the simplex is discrete or continuous. This is a restatement of for example Bertin's definition of a line that encapsulates all points on the continuous line.  By modeling the data topology along with its variablity, this work also provides  a generalization of topology preservation as a deformation retraction from graphic space to data space. By using a functional paradigm, we can deconstruct the graphic to the glyph associated with each data point or even to pieces of a glyph; therefore the renderer has full  flexibility in how to to generate the image, while the graphic to data topology maps provide a way to keep track of which part of the image belongs to which data point.

The toy prototype built using this model validates that is usable for a general purpose visualization tool since it can be iteratively integrated into the existing architecture rather than starting from scratch/ Factoring out glyph formation into assembly functions allows for much more clarity in how the glyphs differ. This prototype demonstrates that this framework can generate the fundemental marks, point (scatter plot), line (line chart), and area (bar chart). Furthermore, the grouped and stacked bar examples demonstrate that this model supports composition of glyphs into more complex graphics. These composite examples also rely on the fiber bundles section base book keeping to keep track of which components contribute to the attributes of the glyph. Implementing this example using a Pandas dataframe demonstrates the ease of incorporating existing widely used data containers rather than requiring users to conform to one stands.  
\subsection{What are the limitations?}
%%%An interesting control problem will be how to keep the text and graphics consistent. If the rendering exchanges the horizontal and vertical axes of a bar chart, the text must also exchange its references to the axes.
\begin{itemize}
\item is at arc level
\item is at one dataset level - needs invariants/composites
\item static (what happens with interactive - but this is just update )
\item is compose(encodes) and ggplot is plot(encodes, glyph)
\item can't handle figerative/semantic equivariance (this actually could maybe be done as extension of fber)
\item explicitely does not care about effectiveness - is domain specific 
\item is only developed against Matplotlib
\item is written against AGG/SVG
\item layout
\item text
\item same as APT, is at gd level so cannot control render 
\item keeping all data in sync if transform is changed in one place (so on multi composed graphs, changing x on does it propopgate - part of Bertin invariance)
\end{itemize}
\subsection{What else needs to be implemented?}
\begin{itemize}
    \item heatmaps/ proof of concept of face simplex
    \item composition algebra
    \item boxplot - multiglyph composition
    \item debatable: non-trivial bundle like data on mobius backend
    \item interactivity 
    \item concurrent artist
    \item numpy wrapper
\end{itemize}
\begin{itemize}
    \item provisional implementation using new architecture in Matplotlib
    \item developer-aimed documentation (API + narrative)
    \item proof of concept 3rd party user-facing library with a biology partner
    \item academic paper about building visualization tools on top of the core architecture
\end{itemize}
Other things that may or may not play a big role would be working out the details of the S->K interaction for dashboards. By this I particularly mean when S is a composite object spanning all the figures shown in the dashboard and a component of K may have different compoents of S mapping into it slicing it in different ways.
\subsection{What are the lessons learned?}
